## Source All Files if still needed:
```{r }
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

library("nimble")
library("coda")
library("splines2")
```

```{r }
source("..\\R_Code\\helpfunctions.r")
# #source("..\\R_Code\\lmm.r")
# #source("..\\R_Code\\Beta_Modell.r")
# source("..\\R_Code\\Beta_Modell_age_spline.r")
# source("..\\R_Code\\Beta_Modell_day_spline.r")
# source("..\\R_Code\\Beta_Modell_age_day_spline.r")
# source("..\\R_Code\\Beta_Modell_day_pen_spline.r")
```

## LMM sofa
```{r }
sofa_lmm <- readRDS("..\\data\\mcmc_res\\sofa_lmm.Rds")

```
```{r }
sofa_lmms <- subset_coda_params(data = sofa_lmm$samples, variable = c("sigma2", "beta0", "beta1", "beta2","sigma2_b0","tau"))
cat("\n the WAIC \n")
sofa_lmm$WAIC$WAIC
summarise_default(sofa_lmms)
```

### Explore predictions
```{r }
y_pred_lmm <- get_params(c("y_pred"),sofa_lmm$samples)
plot(data$sofa, y_pred_lmm)
exists <- !is.na(data$sofa)
cor(data$sofa[exists], y_pred_lmm[exists])

data$sofa_pred <- y_pred_lmm
```


## Beta beta_rand
### some Summary
```{r }
source("helpfunctions.r")
#source("Beta_Modell.r")
#result ist beta_rand
beta_rand <- readRDS("..\\data\\mcmc_res\\beta_rand.rds")

lowe_beta_rand <- readRDS("..\\data\\mcmc_res\\lowe_beta_rand.rds")
lowe_beta_rand$WAIC$WAIC
```

###Test if in general sensitive to changing added value
```{r }
lowe_ai_spline_age <- readRDS("..\\data\\mcmc_res\\lowe_ai_spline_age.rds")
lowe_ai_spline_age$WAIC$WAIC

lowe_ai_day_pen_spline <- readRDS("..\\data\\mcmc_res\\lowe_ai_day_pen_spline.Rds")
lowe_ai_day_pen_spline$WAIC$WAIC

```

```{r }
ai_beta <- subset_coda_params(data = beta_rand$samples, variable = c("beta0", "beta1", "beta2","sigma_b0","phi"))
beta_rand$WAIC$WAIC
summarise_default(ai_beta)
```

### Explore predictions since this one yields the best results, we stick with this model
```{r }
#that got canceled for the moment to save some more room
y_pred_beta <- get_params(c("y_pred"),beta_rand$samples)
data$ai_pred <- y_pred_beta
saveRDS(data, file = "..\\data\\Prep_data_for_ex3.rds")
```
### some Plots

```{r}
source("helpfunctions.r")
beta_means <- get_params(c("beta0", "beta_age", "beta_day","phi"),beta_rand$samples,functions_ = function(x)x)
age_orig <- 55
day_orig <- c(-10:18)
#random <- 0.42
random <- 0.4
get_params("sigma_b0",beta_rand$samples)

beta_day_spline <- function(x){
  mu <- expit(x["beta0"]+x["beta_age"]*age_orig + x["beta_day"]*day_orig)
  phi <- x["phi"]
  a <- mu * phi
  b <- (1 - mu) * phi
  return(qbeta(0.5, a , b))
}
est_values <- (apply(beta_means,1,beta_day_spline))
get_estimates <- apply(est_values,1,quantile,c(0.025,0.5,0.975))

gr1_low <- get_estimates[1,]
gr1_mid <- get_estimates[2,]
gr1_up  <- get_estimates[3,]
plot(data$day, data$ai,  pch = 19, lwd = 1.5,
      xlab = "Days", cex.lab = 1.8,
     ylab = "AI Score", col = data$id)
lines(day_orig,gr1_low, col = 2,lty = 2, lwd = 3)
lines(day_orig,gr1_mid, col = 2, lwd = 3)
lines(day_orig,gr1_up, col = 2,lty = 2, lwd = 3)
# legend("topright", legend = c("SS", "HB"),lwd = 3,
#                lty = 1,col = c(2,3),
#                title = "Groups")


# source("helpfunctions.r")
# #beta 1 age beta2 day
# beta_means <- get_params(c("beta0", "beta1","beta2","sigma_b0","phi"),beta_rand$samples)
# #age and day are transformed therefore also here some transofrmation
# age <- -0
# day <- c(-6,22)
# 
# mu <- expit(beta_means["beta0"]+beta_means["beta1"]*age + beta_means["beta2"]*day)
# phi <- beta_means["phi"]
# a <- mu * phi
# b <- (1 - mu) * phi
# 
# gr1_low <- qbeta(0.025, a , b)
# gr1_mid <- qbeta(0.5, a , b)
# gr1_up <- qbeta(0.975, a , b)
# 
# plot(data$day, data$ai,  pch = 19, lwd = 1.5,
#       xlab = "Days", cex.lab = 1.8,
#      ylab = "Overall survival probability")
# lines(day,gr1_low, col = 2,lty = 2, lwd = 3)
# lines(day,gr1_mid, col = 2, lwd = 3)
# lines(day,gr1_up, col = 2,lty = 2, lwd = 3)
# legend("topright", legend = c("SS", "HB"),lwd = 3,
#                lty = 1,col = c(2,3),
#                title = "Groups")


```
###Compute DIC
```{r}
#beta 1 age beta2 day
data$ai2 <- data$ai + 1e-3
devi <- get_params(c("Deviance"),beta_rand$samples,functions_ = function(x)x)
md <- mean(devi)
# plot(dbeta((1:999) *0.001,1.5,24))
beta_means <- get_params(c("beta0", "beta1","beta2","sigma_b0","phi","b0"),beta_rand$samples)
#age and day are transformed therefore also here some transofrmation
b0_means <- get_params(c("b0"),beta_rand$samples)
b0_subset <- b0_means[data$id]
#plot(density(b0_means))
age <- data$age
day <- data$day
mu <- expit(beta_means["beta0"]+beta_means["beta1"]*data$age + beta_means["beta2"]*data$day+b0_subset)
phi <- beta_means["phi"]
a <- mu * phi
b <- (1 - mu) * phi
pd <- md - (-2*sum(dbeta(data$ai2, a , b,log = T)))
c(pd,md,pd+md)
```
### ppo cpo lpml
```{r}
mcmc_subset <- get_params(c("ppo"),beta_rand$samples,function(x)x)
#cpo (leave one out Prediction)
cpo <- (apply(1/as.matrix(mcmc_subset),2,mean))^-1
icpo <- cpo^-1
#ppo (without leave one out, therefore violates liklihoodprinciple(dont predict with same data))
ppo <- (apply(as.matrix(mcmc_subset),2,mean))
ippo <- ppo^-1

plot(icpo)
plot(ippo)
order(icpo,decreasing = T)[1:7]
data$id[order(icpo,decreasing = T)][1:7]
data$id[order(ippo,decreasing = T)][1:7]
LPML_beta <- mean(log(cpo))
LPML_beta
```

## Beta Age Spline ai_spline_age
```{r}
source("helpfunctions.r")
#source("Beta_Modell_age_spline.r")
ai_spline_age <- readRDS("..\\data\\mcmc_res\\ai_spline_age.Rds")
##result is ai_spline_age
```
### Summary
```{r}
ai_spline_age_subset <- subset_coda_params(c("beta0", "beta_age", "beta2","sigma_b0","phi"),ai_spline_age$samples)
summarise_default(ai_spline_age_subset)
```

```{r}
ai_spline_age$WAIC$WAIC
```

### Plotting
```{r}

#beta 1 age beta2 day
beta_means <- get_params(c("beta0", "beta_age", "beta2","sigma_b0","phi"),ai_spline_age$samples)
#age and day are transformed therefore also here some transofrmation
age <- 22
day <- c(-6:22)
#works tryed it with the true data, if we would make a bigger project do a unit_test
age <- bSpline(age,Boundary.knots =c(min(data$age),max(data$age)) , knots = c(-20,0,20),degree = 3)
mu <- expit(beta_means["beta0"]+c(beta_means[grepl("beta_age", names(beta_means))]%*%t(age)) + beta_means["beta2"]*day)

phi <- beta_means["phi"]
a <- mu * phi
b <- (1 - mu) * phi

gr1_low <- qbeta(0.025, a , b)
gr1_mid <- qbeta(0.5, a , b)
gr1_up <- qbeta(0.975, a , b)

plot(data$day, data$ai,  pch = 19, lwd = 1.5,
      xlab = "Days", cex.lab = 1.8,
     ylab = "Overall survival probability", col = data$id)
lines(day,gr1_low, col = 2,lty = 2, lwd = 3)
lines(day,gr1_mid, col = 2, lwd = 3)
lines(day,gr1_up, col = 2,lty = 2, lwd = 3)
legend("topright", legend = c("SS", "HB"),lwd = 3,
               lty = 1,col = c(2,3),
               title = "Groups")
```
``
### ppo cpo lpml
```{r}
mcmc_subset <- get_params(c("ppo"),ai_spline_age$samples,function(x)x)
#cpo (leave one out Prediction)
cpo <- (apply(1/as.matrix(mcmc_subset),2,mean))^-1
icpo <- cpo^-1
#ppo (without leave one out, therefore violates liklihoodprinciple(dont predict with same data))
ppo <- (apply(as.matrix(mcmc_subset),2,mean))
ippo <- ppo^-1

plot(icpo)
plot(ippo)
order(icpo,decreasing = T)[1:7]
data$id[order(icpo,decreasing = T)][1:7]
data$id[order(ippo,decreasing = T)][1:7]
LPML_ai_spline_age <- mean(log(cpo))
LPML_ai_spline_age

```

## Beta Day Spline
```{r}
source("helpfunctions.r")
#source("Beta_Modell_day_spline.r")

ai_spline_day <- readRDS("..\\data\\mcmc_res\\ai_spline_day.Rds")

#result is ai_spline_age_day
```

### Summary 
```{r }
ai_spline_day_subset <- subset_coda_params(c("beta0", "beta_age", "beta2","sigma_b0","phi"),ai_spline_day$samples)
summarise_default(ai_spline_day_subset)
```
### Plot
```{r}
source("helpfunctions.r")
#beta 1 age beta2 day
#beta_means <- get_params(c("beta0", "beta_age", "beta_day","sigma_b0","phi"),ai_spline_day$samples)
#just the same as subsetting basically
beta_means <- get_params(c("beta0", "beta_age", "beta_day","sigma_b0","phi"),ai_spline_day$samples,functions_ = function(x)x)
#calculate prediction for every set of variables, for chosen x values
#really hand crafted
colnames(beta_means)

age_orig <- 55
day_orig <- c(-10:18)
#var coming from random intercept
#get_params(c("sigma_b0"),ai_spline_day$samples)
random <- 0.42
#works tryed it with the true data, if we would make a bigger project do a unit_test
day <- bSpline(-10:18,Boundary.knots =c(min(data$day),max(data$day)) , knots = c(-4,-2,3),degree = 3)

beta_day_spline <- function(x){
  mu <- expit(2*random + x["beta0"]+x["beta_age"]*age_orig + c(x[grepl("beta_day", names(x))]%*%t(day)))
  phi <- x["phi"]
  a <- mu * phi
  b <- (1 - mu) * phi
  return(qbeta(0.5, a , b))
}
est_values <- (apply(beta_means,1,beta_day_spline))
#now we have all estimates for all sets off data, now calculate there the quantiles
#the objedct has dim (length(day),length(mcmc_sample))
get_estimates <- apply(est_values,1,quantile,c(0.025,0.5,0.975))

gr1_low <- get_estimates[1,]
gr1_mid <- get_estimates[2,]
gr1_up  <- get_estimates[3,]
plot(data$day, data$ai,  pch = 19, lwd = 1.5,
      xlab = "Days", cex.lab = 1.8,
     ylab = "Overall survival probability", col = data$id)
lines(day_orig,gr1_low, col = 2,lty = 2, lwd = 3)
lines(day_orig,gr1_mid, col = 2, lwd = 3)
lines(day_orig,gr1_up, col = 2,lty = 2, lwd = 3)
legend("topright", legend = c("SS", "HB"),lwd = 3,
               lty = 1,col = c(2,3),
               title = "Groups")
# 
# #other method where i just use the quantiles of the total mean of the parameters instead of the quantiles of the estimated values is probably worse
# mu <- expit(beta_means["beta0"]+beta_means["beta_age"]*age_orig + c(beta_means[grepl("beta_day", names(beta_means))]%*%t(day)))
# phi <- beta_means["phi"]
# a <- mu * phi
# b <- (1 - mu) * phi
# 
# gr1_low <- qbeta(0.025, a , b)
# gr1_mid <- qbeta(0.5, a , b)
# gr1_up <- qbeta(0.975, a , b)
# plot(data$day, data$ai,  pch = 19, lwd = 1.5,
#       xlab = "Days", cex.lab = 1.8,
#      ylab = "Overall survival probability", col = data$id)
# lines(day_orig,gr1_low, col = 2,lty = 2, lwd = 3)
# lines(day_orig,gr1_mid, col = 2, lwd = 3)
# lines(day_orig,gr1_up, col = 2,lty = 2, lwd = 3)
# legend("topright", legend = c("SS", "HB"),lwd = 3,
#                lty = 1,col = c(2,3),
#                title = "Groups")
```
###Compute DIC
It misses for the moment b0 and the correct spline specification
```{r}
#beta 1 age beta2 day
devi <- get_params(c("Deviance"),ai_spline_day$samples,functions_ = function(x)x)
md <- mean(devi)

b0_means <- get_params(c("b0"),ai_spline_day$samples)
b0_subset <- b0_means[data$id]

beta_means <- get_params(c("beta0", "beta_age", "beta_day","sigma_b0","phi"),ai_spline_day$samples)

dd_quant <- quantile(data$day,c(0.25,0.5,0.75))
day_spline <- bSpline(data$day, knots = c(dd_quant[1],dd_quant[2],dd_quant[3]),degree = 3)
phi <- beta_means["phi"]
age <- data$age
day <- data$day

a <- mu * phi
b <- (1 - mu) * phi
pd <- md - (-2*sum(dbeta(data$ai2, a , b,log = T)))
c(pd,md,pd+md)
plot(density(b0_means))
```
### ppo cpo lpml
```{r}
mcmc_subset <- get_params(c("ppo"),ai_spline_day$samples,function(x)x)
#cpo (leave one out Prediction)
cpo <- (apply(1/as.matrix(mcmc_subset),2,mean))^-1
icpo <- cpo^-1
#ppo (without leave one out, therefore violates liklihoodprinciple(dont predict with same data))
ppo <- (apply(as.matrix(mcmc_subset),2,mean))
ippo <- ppo^-1

plot(icpo)
plot(ippo)
order(icpo,decreasing = T)[1:7]
data$id[order(icpo,decreasing = T)][1:7]
data$id[order(ippo,decreasing = T)][1:7]
LPML_ai_spline_day <- mean(log(cpo))
LPML_ai_spline_day

```
## Beta Day Age Spline ai_spline_age_day
```{r}
#source("Beta_Modell_age_day_spline.r")

ai_spline_age_day <- readRDS("..\\data\\mcmc_res\\ai_spline_age_day.Rds")
#result is ai_spline_age_day
```
###Summary
```{r }
ai_spline_age_day_subset <- subset_coda_params(c("beta0", "beta_age", "beta2","sigma_b0","phi"),ai_spline_age_day$samples)
summarise_default(ai_spline_age_day_subset)
```
```{r }
ai_spline_age_day$WAIC$WAIC
```


###Plotting
```{r}
source("helpfunctions.r")
#beta 1 age beta2 day
beta_means <- get_params(c("beta0", "beta_age", "beta_day","sigma_b0","phi"),ai_spline_age_day$samples)
#age and day are transformed therefore also here some transofrmation
age_orig <- 22
day_orig <- c(-6:22)
#works tryed it with the true data, if we would make a bigger project do a unit_test
age <- bSpline(age_orig,Boundary.knots =c(min(data$age),max(data$age)) , knots = c(-20,0,20),degree = 3)
day <- bSpline(day_orig,Boundary.knots =c(min(data$day),max(data$day)) , knots = c(-4,-2,3),degree = 3)

mu <- expit(beta_means["beta0"]+c(beta_means[grepl("beta_age", names(beta_means))]%*%t(age)) + c(beta_means[grepl("beta_day", names(beta_means))]%*%t(day)))
phi <- beta_means["phi"]
a <- mu * phi
b <- (1 - mu) * phi


gr1_low <- qbeta(0.025, a , b)
gr1_mid <- qbeta(0.5, a , b)
gr1_up <- qbeta(0.975, a , b)
plot(data$day, data$ai,  pch = 19, lwd = 1.5,
      xlab = "Days", cex.lab = 1.8,
     ylab = "Overall survival probability", col = data$id)
lines(day_orig,gr1_low, col = 2,lty = 2, lwd = 3)
lines(day_orig,gr1_mid, col = 2, lwd = 3)
lines(day_orig,gr1_up, col = 2,lty = 2, lwd = 3)
legend("topright", legend = c("SS", "HB"),lwd = 3,
               lty = 1,col = c(2,3),
               title = "Groups")
```
### ppo cpo lpml
```{r}
mcmc_subset <- get_params(c("ppo"),ai_spline_age_day$samples,function(x)x)
#cpo (leave one out Prediction)
cpo <- (apply(1/as.matrix(mcmc_subset),2,mean))^-1
icpo <- cpo^-1
#ppo (without leave one out, therefore violates liklihoodprinciple(dont predict with same data))
ppo <- (apply(as.matrix(mcmc_subset),2,mean))
ippo <- ppo^-1

plot(icpo)
plot(ippo)
order(icpo,decreasing = T)[1:7]
data$id[order(icpo,decreasing = T)][1:7]
data$id[order(ippo,decreasing = T)][1:7]
LPML_ai_spline_age_day <- mean(log(cpo))
LPML_ai_spline_age_day

```


## Beta P-Bsplines Day
```{r}
#source("..\\R_code\\Beta_Modell_day_pen_spline.r")
ai_pspline_day <-readRDS("..\\data\\mcmc_res\\day_pen_spline.Rds")
#result is ai_spline_age_day
```

### Summary 
```{r }
ai_pspline_day_subset <- subset_coda_params(c("beta0", "beta_age", "beta2","sigma_b0","phi"),ai_pspline_day$samples)
summarise_default(ai_pspline_day_subset)
```

### with splines a for day a plot
```{r}
# source("helpfunctions.r")
# #beta 1 age beta2 day
# beta_means <- get_params(c("beta0", "beta_age", "beta_day","sigma_b0","phi"),ai_pspline_day$samples)
# #age and day are transformed therefore also here some transofrmation
# #works tryed it with the true data, if we would make a bigger project do a unit_test
# day <- bSpline(day_orig,degree = 3,df = 20)
# age_orig <- 22
# day_orig <- seq(-15,22,length= ncol(day))
# 
# mu <- expit(beta_means["beta0"]+beta_means["beta_age"]*age_orig + c(beta_means[grepl("beta_day", names(beta_means))]%*%t(day)))
# phi <- beta_means["phi"]
# a <- mu * phi
# b <- (1 - mu) * phi
# 
# 
# gr1_low <- qbeta(0.025, a , b)
# gr1_mid <- qbeta(0.5, a , b)
# gr1_up <- qbeta(0.975, a , b)
# plot(data$day, data$ai,  pch = 19, lwd = 1.5,
#       xlab = "Days", cex.lab = 1.8,
#      ylab = "Overall survival probability", col = data$id)
# lines(day_orig,gr1_low, col = 2,lty = 2, lwd = 3)
# lines(day_orig,gr1_mid, col = 2, lwd = 3)
# lines(day_orig,gr1_up, col = 2,lty = 2, lwd = 3)
# legend("topright", legend = c("SS", "HB"),lwd = 3,
#                lty = 1,col = c(2,3),
#                title = "Groups")
```
### ppo cpo lpml
```{r}
mcmc_subset <- get_params(c("ppo"),ai_pspline_day$samples,function(x)x)
#cpo (leave one out Prediction)
cpo <- (apply(1/as.matrix(mcmc_subset),2,mean))^-1
icpo <- cpo^-1
#ppo (without leave one out, therefore violates liklihoodprinciple(dont predict with same data))
ppo <- (apply(as.matrix(mcmc_subset),2,mean))
ippo <- ppo^-1

plot(icpo)
plot(ippo)
order(icpo,decreasing = T)[1:7]
data$id[order(icpo,decreasing = T)][1:7]
data$id[order(ippo,decreasing = T)][1:7]
LPML_ai_pspline_day <- mean(log(cpo))
LPML_ai_pspline_day

```


## Beta P-Bsplines Age
```{r}
#source("..\\R_code\\Beta_Modell_day_pen_spline.r")
ai_pspline_age <-readRDS("..\\data\\mcmc_res\\ai_age_pen_spline.Rds")
#result is ai_spline_age_day
```



### Summary 
```{r }
ai_pspline_age_subset <- subset_coda_params(c("beta0", "beta_age", "beta2","sigma_b0","phi"),ai_pspline_age$samples)
summarise_default(ai_pspline_age_subset)
```
### ppo cpo lpml
```{r}
mcmc_subset <- get_params(c("ppo"),ai_pspline_age$samples,function(x)x)
#cpo (leave one out Prediction)
cpo <- (apply(1/as.matrix(mcmc_subset),2,mean))^-1
icpo <- cpo^-1
#ppo (without leave one out, therefore violates liklihoodprinciple(dont predict with same data))
ppo <- (apply(as.matrix(mcmc_subset),2,mean))
ippo <- ppo^-1

plot(icpo)
plot(ippo)
order(icpo,decreasing = T)[1:7]
data$id[order(icpo,decreasing = T)][1:7]
data$id[order(ippo,decreasing = T)][1:7]
LPML_ai_pspline_age <- mean(log(cpo))
LPML_ai_pspline_age

```
## WAIC comparison
```{r}
cat("no spline WAIC\n")
beta_rand$WAIC$WAIC
cat("\n spline age WAIC\n")
ai_spline_age$WAIC$WAIC
cat("\n spline day WAIC\n")
ai_spline_day$WAIC$WAIC
cat("\n spline day, age WAIC\n")
ai_spline_age_day$WAIC$WAIC
cat("\n pspline day WAIC\n")
ai_pspline_day$WAIC$WAIC
cat("\n pspline age WAIC\n")
ai_pspline_age$WAIC$WAIC
```
## LPML (Bayes Factor)
```{r}
cat("no spline WAIC\n")
LPML_beta
cat("\n spline age WAIC\n")
LPML_ai_spline_age
cat("\n spline day WAIC\n")
LPML_ai_spline_day
cat("\n spline day, age WAIC\n")
LPML_ai_spline_age_day
cat("\n pspline day WAIC\n")
LPML_ai_pspline_day
cat("\n pspline age WAIC\n")
LPML_ai_pspline_age
```